Spatio-Temporal Modeling and Simulation

### 2024-06-12 report/Report.tex
log odds boi

\section{Odds Modeling}

@@

$$P'(v_i|\eta_i) = \frac{\psi_i}{\sum_k \psi_k}$$

These models will be known as odds models as they predict the "odds for" each outcome $v_k$ given the information contained in $\eta_k$.

@@

Then our proposed probabilities are:

$$P_i'(v_{jk}|\eta_{jk})=\frac{\phi_{jki}}{\sum_p \phi_{jpi}}$$

Then if the $k$ selected per decision $D_j$ is given by $s_j$ we want to maximize:

@@

$$\sum_j \ln{\left[ P_i'(v_{js_j} | \eta_{js_j}) \right]}= \sum_j \ln{\left[ \frac{\phi_{js_ji}}{\sum_p \phi_{jpi}} \right]}$$

@@

$$\partial_{\phi_{js_ji}}\ln{\mathcal{L}}=\frac{1}{P_i'(v_{js_j} | \eta_{js_j})}\frac{\sum_p \phi_{jpi}-\phi_{js_ji}}{\left( \sum_p \phi_{jpi} \right)^2}=\frac{1}{\sum_p \phi_{jpi}}\frac{\left(1-P_i'(v_{js_j} | \eta_{js_j}) \right)}{P_i'(v_{js_j} | \eta_{js_j})}$$

$$\partial_{\phi_{j\not{s_j}i}}\ln{\mathcal{L}}=\frac{1}{P_i'(v_{js_j} | \eta_{js_j})}\frac{-\phi_{js_ji}}{\left( \sum_p \phi_{jpi} \right)^2} = \frac{-1 }{\sum_p \phi_{jpi}}$$

@@

$$\partial_Z \ln{\mathcal{L}}=\sum_{\phi_{js_ji} \in Z}\frac{1}{\sum_p \phi_{jpi}}\frac{\left(1-P_i'(v_{js_j} | \eta_{js_j}) \right)}{P_i'(v_{js_j} | \eta_{js_j})}-\sum_{\phi_{j\not{s_j}i}\in Z}\frac{1 }{\sum_p \phi_{jpi}}$$

@@

\begin{procedure}{Fitting an Odds Model}
\begin{enumerate}
\item Collect decisions $D_j$ and corresponding features $\eta_{jk}$, options $v_{jk}$, and selection $s_j$. 
\item Make an initial guess $\psi_{jk0} = 1$.
\item Fit $\hat{G}_i$ on the pairs of $\psi_{jk(i-1)},\eta_{jk}$ using MSE to produce the $\phi_{jki}$.
\item Generate the $u_{jki} = \alpha_i \partial_{\phi_{jki}} \ln{\mathcal{L}}$ and produce a new set of $\psi_{jki}$.
\item Repeat 3 and 4 until, varying $\alpha_i$ until convergence.

$$\partial_{\phi_{js_ji}}\ln{\mathcal{L}}=\frac{1}{\sum_p \phi_{jpi}}\frac{\left(1-P_i'(v_{js_j} | \eta_{js_j}) \right)}{P_i'(v_{js_j} | \eta_{js_j})}$$

$$\partial_{\phi_{j\not{s_j}i}}\ln{\mathcal{L}}=\frac{-1 }{\sum_p \phi_{jpi}}$$ 
\end{enumerate}

>>>>

\section{Log Odds Modeling}

@@

$$P'(v_i|\eta_i) = \frac{e^{\psi_i}}{\sum_k e^{\psi_k}}$$

These models will be known as log odds models as they predict the "log odds for" each outcome $v_k$ given the information contained in $\eta_k$.

@@

Our probability is therefore:

$$P_i'(v_{jk}|\eta_{jk})=\frac{e^{\phi_{jki}}}{\sum_p e^{\phi_{jpi}}}$$

Now if the $k$ selected per decision $D_j$ is given by $s_j$ we want to maximize:

@@

$$\sum_j \ln{\left[ P_i'(v_{js_j} | \eta_{js_j}) \right]}= \sum_j \ln{\left[ \frac{e^{\phi_{js_ji}}}{\sum_p e^{\phi_{jpi}}} \right]}$$

@@

$$\partial_{\phi_{js_ji}}\ln{\mathcal{L}}
=
\frac{1}{P_i'(v_{js_j} | \eta_{js_j})}\frac{\sum_p e^{\phi_{jpi}}-e^{\phi_{js_ji}}}{\left( \sum_p e^{\phi_{jpi}} \right)^2} e^{\phi_{js_ji}}
=
\frac{1}{\sum_p e^{\phi_{jpi}}}\frac{\left(1-P_i'(v_{js_j} | \eta_{js_j}) \right)}{P_i'(v_{js_j} | \eta_{js_j})} e^{\phi_{js_ji}}$$

$$\partial_{\phi_{j\not{s_j}i}}\ln{\mathcal{L}}
=
\frac{1}{P_i'(v_{js_j} | \eta_{js_j})}\frac{-e^{\phi_{js_ji}}}{\left( \sum_p e^{\phi_{jpi}} \right)^2} e^{\phi_{j\not{s_j}i}}
= 
\frac{-1 }{\sum_p e^{\phi_{jpi}}}e^{\phi_{j\not{s_j}i}}$$

@@

$$\partial_Z \ln{\mathcal{L}}=\sum_{\phi_{js_ji} \in Z}\frac{1}{\sum_p \phi_{jpi}}\frac{\left(1-P_i'(v_{js_j} | \eta_{js_j}) \right)}{P_i'(v_{js_j} | \eta_{js_j})}e^{\phi_{js_ji}}-\sum_{\phi_{j\not{s_j}i}\in Z}\frac{1 }{\sum_p \phi_{jpi}}e^{\phi_{j\not{s_j}i}}$$

@@

\begin{procedure}{Fitting a Log Odds Model}
\begin{enumerate}
\item Collect decisions $D_j$ and corresponding features $\eta_{jk}$, options $v_{jk}$, and selection $s_j$. 
\item Make an initial guess $\psi_{jk0} = 0$.
\item Fit $\hat{G}_i$ on the pairs of $\psi_{jk(i-1)},\eta_{jk}$ using MSE to produce the $\phi_{jki}$.
\item Generate the $u_{jki} = \alpha_i \partial_{\phi_{jki}} \ln{\mathcal{L}}$ and produce a new set of $\psi_{jki}$.
\item Repeat 3 and 4 until, varying $\alpha_i$ until convergence.

$$\partial_{\phi_{js_ji}}\ln{\mathcal{L}}=\frac{1}{\sum_p e^{\phi_{jpi}}}\frac{\left(1-P_i'(v_{js_j} | \eta_{js_j}) \right)}{P_i'(v_{js_j} | \eta_{js_j})}e^{\phi_{js_ji}}$$

$$\partial_{\phi_{j\not{s_j}i}}\ln{\mathcal{L}}=\frac{-1 }{\sum_p e^{\phi_{jpi}}}e^{\phi_{j\not{s_j}i}}$$ 
\end{enumerate}
